---
title: "Wildfire"
subtitle: "Part C: Monte Carlo"
author: "YOUR NAME"
date: today
format: html
---

# Introduction

Idaho is well known for its wildfires. They play a ecological role, shaping our forests and grasslands while presenting significant hazards to communities and resources. In recent years, the Idaho landscape has experienced frequent and sometimes severe wildfire seasons due to both natural and human activity. While fires are part of the natural ecosystem, they have become more destructive as the climate warms. Wildfire preparedness and management is a continuing challenge for the state.

Wildfire managers need tools that show where fires are more likely and how intense they could become. The U.S. Forest Service’s Wildfire Hazard Potential (WHP) is one such tool: a nationwide raster layer where each grid cell stores a unitless index of relative wildfire hazard (higher value = higher hazard).

How far could a new wildfire run in 8 hours? How many new wild fire ignitions will grow larger than 10 km (straight line) in 8 hours?


# `rlnorm` + `replicate()` 

❓ Question e: In a season with 450 new ignitions **statewide**, what is the probability we see at least 50 fast-growing events (≥10 km in 8 hours) when ignitions can occur in any hazard category according to their statewide area weights?

## Create category labels

**TASK 1**

🎯 **Goal.** ____________________________

```{r}
# ⚡ Test: vector of category labels
____  # create vector <- c( ... ) using the Category names from Table 1
```

```{r}
# ✅  Verify: inspect labels

____ # display labels
```

🗣 Comment (TODO).


## Create category probabilities

**TASK 2**

🎯 **Goal.** ____________________________

```{r}
# ⚡ Test: vector of category probabilities
____ # create vector <- c( ... ) using the Probability column from Table 1
```

```{r}
# ✅  Verify: : confirm they form a valid distribution
____ # display vector
____ # check that sum() is 1
```

## Create lognormal parameter vectors

**TASK 3**

🎯 **Goal.** ____________________________


```{r}
# ⚡ Test: vectors of lognormal parameters
____ # create vector 1 <- c( ...) from the Mean log column
____ # create vector 2 <- c( ...) from the SD log column
```

```{r}
# ✅  Verify: inspect and compare
____ # display vector 1
____ # display vector 2
____ # check length of labels vector, with vector 1 and 2.
```

🗣 Comment (TODO).


## Set global simulation constants

**TASK 4**

🎯 **Goal.** ____________________________

```{r}
# ⚡ Test: number of ignitions, seasons, and fast threshold
____ # set N to the number of ignitions per season 
____ # set S to the number of simulated seasons
____ # set k to the fast-run distance threshold in km 
```

```{r}
# ✅  Verify: inspect constants
____ # display N
____ # display S
____ # display k
```

## Sample categories for one season

**TASK 5**

🎯 **Goal.** ____________________________


```{r}
# ⚡ Test: sample category indices for one season
____ # create idx using sample.int() (or sample()) with size = N, prob = prob
```


```{r}
# ✅  Verify: examine realized category mix

____ # inspect the first few indices
____ # use table() to see counts per category
```

🗣 Comment (TODO).


## Draw distances for one season

**TASK 6**

🎯 **Goal.** ____________________________


```{r}
# ⚡ Test: simulate distances for one season
____ # store d, use rlnorm() with n = N, meanlog = meanlog[idx], sdlog = sdlog[idx]
```

```{r}
# ✅  Verify: examine distances

____ # show head()
____ # summarize d 
```

🗣 Comment (TODO).


## Count fast runs for one season

**TASK 7**

🎯 **Goal.** ____________________________


```{r}
# ⚡ Test: count fast runs in one season
____ # create fast_flags_one as a logical vector using d >= k
____ # compute fast_count_one using sum() of fast_flags_one
```
```{r}
# ✅  Verify: inspect count and proportion
____ # display fast_count_one
____ # optionally compute mean() of fast_flags_one to get proportion
```

🗣 Comment (TODO).


## Bundle one-season logic in a single block

**TASK 8**


🎯 **Goal.** ____________________________

```{r}
# ⚡ Test: bundled one-season simulation returning a single fast count
____ <- { #store vector and start block
____ # sample idx for one season
____ # simulate distances d using meanlog, sdlog with brackets and idx
____ # return the count of d values that are greater than or equal to k
} #end block
```
```{r}
# ✅  Verify: confirm the block works
____ # display fast_count_one
```

🗣 Comment (TODO).



## Simulate many seasons

**TASK 9**

🎯 **Goal.** ____________________________

```{r}
# ⚡ Test: many-season simulation with replicate()
set.seed(2026)
____ # store resulting vector, use replicate() 
____ # with n = number of seasons
____ # and expr = { ... one-season block ... } as before
)
```
```{r}
# ⚡ Test:  Answer Question
# summarize seasonal fast counts and estimate probabilities
____ # inspect summary() of counts_fast_mix
____ # compute mean() for expected fast runs
____ # compute mean() for Monte Carlo estimate of P(T >= 50)  
```

🗣 Comment (TODO).


# Visualize the distribution of seasonal fast counts

**TASK 10**

🎯 **Goal.** ____________________________

```{r}
# ⚡ Test: histogram of seasonal fast counts

____ # load ggplot2 if you choose to use it
____ # create a data frame with a column fast_count = counts_fast_mix

____ # build a histogram of fast_count (e.g., ggplot(...) + geom_histogram(...))
____ # Add a vertical line at the threshold of 50 fast fires
```

```{r}
# ✅  Verify: inspect histogram
____ # display the histogram
```

🗣 Comment (TODO).















