---
title: "Wildfire"
subtitle: "Part A: Explore and Wrangle"
author: "YOUR NAME"
date: today
format: html
---

# Introduction

Idaho is well known for its wildfires. They play a ecological role, shaping our forests and grasslands while presenting significant hazards to communities and resources. In recent years, the Idaho landscape has experienced frequent and sometimes severe wildfire seasons due to both natural and human activity. While fires are part of the natural ecosystem, they have become more destructive as the climate warms. Wildfire preparedness and management is a continuing challenge for the state.

Wildfire managers need tools that show where fires are more likely and how intense they could become. The U.S. Forest Service’s Wildfire Hazard Potential (WHP) is one such tool: a nationwide raster layer where each grid cell stores a unitless index of relative wildfire hazard (higher value = higher hazard).

How far could a new wildfire run in 8 hours? How many new wild fire ignitions will grow larger than 10 km (straight line) in 8 hours?

# Explore and Wrangle

The data analyst's first step is to load the data and explore and modify it. 

## Library and Load Data

**Task 1**

🎯 **Goal.** ____________________________

```{r}
#| message: false
#| warning: false
#⚡ Code
____ # 1. e.g., load any needed package(s)
____ # 2. e.g., read the CSV into county_raw
```


```{r}
#| message: false
#| warning: false
#⚡ Code
county_raw <- readr::read_csv("_______________________")
```


```{r, eval=FALSE}
# ✅  Verify 
dplyr::glimpse(county_raw)

nrow(county_raw)
```

🗣 **Comment**  ____________________________


## Audit Check

**Task 2**

🎯 **Goal.** ____________________________

```{r}
# ⚡ Code: summarize key audit quantities for county_raw
____ # 1/2. assign county_summary and use county_raw
____ # 1. compute n_rows
____ # 1. compute n_na_whp
____ # 1. compute min_area
____ # 1. compute max_area

# ✅ Verify

knitr::kable(county_summary)
```

🗣 **Comment**  ____________________________

## Compute WHP quantile edges

**Task 3**

🎯 **Goal.** ____________________________

```{r}
# ⚡ Code: Base R stats::quantile 
____ # 2. store WHP quantile edges in object e
____ # 1. call quantile() on county_raw's whp column
____ # 1. specify the probs vector
____ # 3. include remove NA from calculation

```
```{r}
# ✅  Verify
e
```

🗣 **Comment**  ____________________________

## Category Label

**Task 4**

🎯 **Goal.** ____________________________

```{r}
# ⚡ Code:  dplyr::mutate + dplyr::case_when
____ # 1. assign county_cats and pipe county_raw into it 
____ # 2. use mutate() to create Category 
____ # 3. with case_when() and e[ ] bounds
____ # 4. convert Category to an ordered factor with the specified levels
```

```{r}
# ✅  Verify: confirm Category looks right
____ # e.g., count() or table()
____ # optional: str() or glimpse()
```

🗣 **Comment**  ____________________________


## Summary Table

**Task 5**

🎯 **Goal.** ____________________________


```{r}
# ⚡ Code: dplyr::group_by + dplyr::summarize + dplyr::mutate
____ # 1. store prob_tbl and pipe county_cats into it
____ # 2. group_by()
____ # 3. summarize()
____ # 4. mutate())
____ # 5. select()
```

```{r}
# ✅  Verify: validate prob_tbl
____ # print prob_tbl
____ # check that probabilities sum to ~1
```

🗣 **Comment**  ____________________________
